
---
title: "Using Custom PNG Images in ggplot2"
author: "Ehsan Zangene"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
# BiocManager::install("ggimage")
library(ggimage)
```

## 🐧 Custom Shapes with PNG Files in ggplot2

This example demonstrates how to use **custom PNG images** to replace default shapes for a variable like `species` in the `penguins` dataset.

### Step 1: Load and Preview the Data

```{r}
source("../peng_data_prep.R")
penguins_clean <- penguins_clean |> 
  drop_na(species, body_mass_g)

head(penguins_clean)
```

### Step 2: Assign an Image Path for Each Species

Ensure you have 3 PNG images (small icons) in an "img" directory in your R project root folder:

- `adelie.png`
- `chinstrap.png`
- `gentoo.png`

```{r}
penguins_img <- penguins_clean |> 
  mutate(image = case_when(
    species == "Adelie" ~ paste0("./Untitled/adelie","_2nd.png"),
    species == "Chinstrap" ~ paste0("./Untitled/chinstrap","_2nd.png"),
    species == "Gentoo" ~ paste0("./Untitled/gentoo","_2nd.png")
  ))

penguins_img
```

### Step 3: Create the Plot Using ggimage

```{r}
ggplot(penguins_img, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_image(aes(image = image), size = 0.05) +
  theme_minimal() +
  labs(
    title = "Penguins: Body Mass vs Flipper Length",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  )

```

### Notes

- The PNG files must be present in the path `img/` relative to your Rmd file.
- Use small and simple PNG files (~32x32 pixels) for better performance and clarity.

### Optional: Save as PNG or PDF

```{r, eval=FALSE}
ggsave("penguins_custom_image_plot.png", width = 8, height = 5)
```

```{r}
library(ggrepel)
penguins_img %>% 
  group_by(species) %>% 
  summarise(count = n(), mean_bil = mean(flipper_length_mm))

ggplot(penguins_img, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  # geom_point(aes(alpha = 0.6), size = 1) +  # Transparent image
  # geom_jitter() +
  geom_image(aes(image = image), size = 0.05) +  # Transparent image
  geom_text_repel(aes(label = species, color = species), size = 3, max.overlaps = 15) +
  # geom_text(aes(label = species, color = species), size = 3, max.overlaps = 20) +
  scale_alpha_identity() +  # This ensures alpha is not scaled to a legend
  scale_color_manual(
    values = c("Adelie" = "#FF8C00", "Gentoo" = "#159090", "Chinstrap" = "#A034F0"),
    name = "Species"
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4)),
    alpha = "none"  # Hide alpha from legend
  ) +
  labs(
    title = "Penguins: Body Mass vs Flipper Length",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  ) +
  theme_minimal(base_size = 5) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    plot.title = element_text(face = "bold")
  )
```

