---
title: "Beautiful R Markdown Report"
author: "Ehsan Zangene"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: tango
  pdf_document:
    toc: true
    number_sections: true
    latex_engine: xelatex
    keep_tex: true
fontsize: 8pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE, 
  fig.align = "center", fig.width = 4, fig.height = 3, dpi = 300
)
library(tidyverse)
library(palmerpenguins)
library(plotly)
```

# ðŸ“˜ Introduction

This R Markdown file demonstrates how to create aesthetically pleasing reports in both **HTML** and **PDF** formats using different chunk options and themes.

---

# ðŸ› ï¸ Chunk Options for Knitting

## ðŸ” Showing results without showing code

```{r no-code, echo=FALSE}
summary(penguins)
```

## ðŸ”§ Hide messages/warnings but show the code

```{r suppress-messages}
library(dplyr)
penguins |> 
  filter(!is.na(flipper_length_mm)) |> 
  group_by(species) |> 
  summarise(avg_flipper = mean(flipper_length_mm, na.rm = TRUE))
```

## ðŸ“Š Include a Plot

```{r penguin-plot}
p <- ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_density(alpha = 0.6) +
  theme_minimal() +
  labs(
    title = "Flipper Length Density by Species",
    x = "Flipper Length (mm)",
    fill = "Species"
  )

ggplotly(p)
```

---

# ðŸŽ¨ Exporting the Report

To knit into:
- **HTML**: Click `Knit` > `Knit to HTML`.
- **PDF**: Install TinyTeX with `tinytex::install_tinytex()` if not already installed, then `Knit to PDF`.

---

# ðŸ“Œ Tips for Aesthetics

- Use `theme_minimal()` or `theme_classic()` in plots.
- Set chunk options like `fig.align = "center"`, `dpi = 300`, `fig.width`, and `fig.height` for better control.
- Customize the YAML header with `theme`, `highlight`, and `toc`.

---

# âœ… Conclusion

R Markdown helps combine code, analysis, and narrative seamlessly. The formatting options enhance the clarity and beauty of your reports.
