
---
title: "Comprehensive Introduction to R Shiny Web App Development"
author: "Ehsan Zangene"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(shiny)
library(tidyverse)
library(palmerpenguins)
library(plotly)
```

# ðŸ§  Introduction

**Shiny** is an R package that makes it easy to build **interactive web applications** straight from R. It is especially powerful for building data dashboards, exploratory tools, and teaching applications.

---

# ðŸ”§ Shiny App Structure

A basic Shiny app has two components:
- **UI**: User Interface
- **Server**: Backend logic

Here's a minimal example:

```{r, eval=FALSE}
library(shiny)

ui <- fluidPage(
  titlePanel("Hello Shiny!"),
  sidebarLayout(
    sidebarPanel(sliderInput("obs", "Number of observations:", 1, 100, 50)),
    mainPanel(plotOutput("distPlot"))
  )
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
    hist(rnorm(input$obs))
  })
}

shinyApp(ui = ui, server = server)
```

---

# ðŸ“ˆ Interactive Data Visualization

You can use **Plotly** and **ggplot2** for interactive and beautiful graphics.

```{r, eval=FALSE}
ui <- fluidPage(
  titlePanel("Interactive Penguin Plot"),
  sidebarLayout(
    sidebarPanel(selectInput("xvar", "X-axis:", choices = names(penguins), selected = "bill_length_mm"),
                 selectInput("yvar", "Y-axis:", choices = names(penguins), selected = "body_mass_g")),
    mainPanel(plotlyOutput("pengPlot"))
  )
)

server <- function(input, output) {
  output$pengPlot <- renderPlotly({
    ggplot(penguins, aes_string(x = input$xvar, y = input$yvar, color = "species")) +
      geom_point() +
      theme_minimal()
  })
}

shinyApp(ui, server)
```

---

# âš™ï¸ Reactive Inputs and Filtering

```{r, eval=FALSE}
ui <- fluidPage(
  titlePanel("Penguin Explorer"),
  sidebarLayout(
    sidebarPanel(
      checkboxGroupInput("speciesInput", "Choose species:", choices = unique(na.omit(penguins$species)), selected = "Adelie")
    ),
    mainPanel(
      tableOutput("filteredTable")
    )
  )
)

server <- function(input, output) {
  output$filteredTable <- renderTable({
    penguins %>%
      filter(species %in% input$speciesInput)
  })
}

shinyApp(ui, server)
```

---

# ðŸ“Š Combining ggplot2 and Interactivity with `plotly`

```{r, eval=FALSE}
ui <- fluidPage(
  titlePanel("Interactive ggplot2 with plotly"),
  mainPanel(
    plotlyOutput("plot1")
  )
)

server <- function(input, output) {
  output$plot1 <- renderPlotly({
    p <- ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
      geom_point(size = 2) +
      theme_minimal()
    ggplotly(p)
  })
}

shinyApp(ui, server)
```

---

# ðŸ“Ž Conclusion

With **Shiny**, you can build powerful and engaging web apps to explore and visualize data interactively. Use `reactive()` functions to capture changes in inputs, and tools like `plotly` to make your visualizations dynamic.

