
---
title: "Day 3 - Interactive Visualizations and Reporting"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(shiny)
```

# Session 1: Static to Interactive with plotly (2h)

- `ggplotly()` from static ggplot2 plots
- Interactive legends, zoom, tooltips

```{r}
p <- ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species))
ggplotly(p)
```

# Session 2: Interactive Dashboards with Shiny (3h)

- Shiny UI and Server basics
- Inputs (selectInput, sliderInput), Outputs (renderPlot, renderTable)
- Building a reactive dashboard

```{r, eval=FALSE}
shinyApp(
  ui = fluidPage(
    selectInput("species", "Select Species", choices = unique(penguins$species)),
    plotOutput("plot")
  ),
  server = function(input, output) {
    output$plot <- renderPlot({
      ggplot(filter(penguins, species == input$species), aes(x = body_mass_g)) +
        geom_histogram()
    })
  }
)
```

# Exercises (2h)

- Build a dashboard for penguin dataset exploration
- Add reactivity and custom themes

# Session 3: R Markdown and HTML Reporting (2h)

- Create reports with `rmarkdown::render()`
- Embed plots, tables, and parameterized reports
- Export HTML and PDF

# Session 4: Reproducibility and GitHub (1h)

- Version control with git
- Using GitHub for collaboration
- Best practices: scripts, reports, documentation

