
---
title: "Using Custom PNG Images in ggplot2"
author: "Ehsan Zangene"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
# BiocManager::install("ggimage")
library(ggimage)
```

## 🐧 Custom Shapes with PNG Files in ggplot2

This example demonstrates how to use **custom PNG images** to replace default shapes for a variable like `species` in the `penguins` dataset.

### Step 1: Load and Preview the Data

```{r}
penguins_clean <- penguins |> 
  drop_na(species, body_mass_g)

head(penguins_clean)
```

### Step 2: Assign an Image Path for Each Species

Ensure you have 3 PNG images (small icons) in an "img" directory in your R project root folder:

- `adelie.png`
- `chinstrap.png`
- `gentoo.png`

```{r}
penguins_img <- penguins_clean |> 
  mutate(image = case_when(
    species == "Adelie" ~ "./img/adelie.png",
    species == "Chinstrap" ~ "./img/chinstrap.png",
    species == "Gentoo" ~ "./img/gentoo.png"
  ))
```

### Step 3: Create the Plot Using ggimage

```{r}
ggplot(penguins_img, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_image(aes(image = image), size = 0.05) +
  theme_minimal() +
  labs(
    title = "Penguins: Body Mass vs Flipper Length",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  )

ggplot(penguins_img, aes(x = flipper_length_mm, y = body_mass_g)) +
  geom_image(aes(image = image), size = 0.05) +
  labs(
    title = "Penguins: Body Mass vs Flipper Length",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  ) +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor = element_line(color = "gray95"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    plot.title = element_text(face = "bold")
  )

```

### Notes

- The PNG files must be present in the path `img/` relative to your Rmd file.
- Use small and simple PNG files (~32x32 pixels) for better performance and clarity.

### Optional: Save as PNG or PDF

```{r, eval=FALSE}
ggsave("penguins_custom_image_plot.png", width = 8, height = 5)
```
